LAI Workflow Documentation

### 2021-02-06 LAI.Rmd
Reformatted adn updated

2. Gap-filled species-level mean leaf lifespan, with genus-level mean. If genus-level not available, then family-level mean. When neither were available: (1) for deciduous species, used mean lifespan at PNM site, (2) for evergreen species, used a relationship of leaf lifespan (LL) with LMA. This LMA-based relationship was based on data at San-Lorenzo (FTS) for canopy leaves and for species with sufficient sample sizes (n_lifespan > 100) (Figure \@ref(fig:ll-lma)). Per Joe, BCI climate is more similar to San Lorenzo, while PNM site is much drier. More importantly, the BCI 50-ha plot and San Lorenzo support old-growth forests. PNM supports much younger, secondary forest (age unknown, perhaps 100 years). So, used the leaf lifespan-LMA relationship for San Lorenzo. This was used to predict leaf lifespan data for species without leaf lifespan data based on their LMA, after gap-filling species LMA data:
    Gap-filled LMALAM_AVD with the following sequence of priority: 
    LMALAM_AVD relationship with LMADISC_AVD > LMALAM_AVD rel. with LMALEAF_AVD > Genus-level mean LMALAM_AVD (gap-filled with LMADISC_AVD/LMALEAF_AVD) > Family-level mean LMALAM_AVD (gap-filled with LMADISC_AVD/LMALEAF_AVD)

@@

fill.day.gaps <- function(df) {
  df.1 <- df %>% arrange(date) %>%
    # for the first census assume leaf fall is collected over the past week
    mutate(n.days = c(7, as.numeric(date - lag(date))[-1])) %>%
    #  leaf_fall rate per day for the census interval is stored on the day of the census:
    mutate(leaf_gm_rate = leaf_gm_per_m2/n.days,
           # approx() will interpolate this forward until the next census,
           # we want it to interpolate this value backward over the past census interval,
           # so just for computation purposes, leading this value
           leaf_gm_rate_lead = lead(leaf_gm_rate, n = 1))
  # Gap-filling dates
  full.date.df <- data.frame(date = seq(from = min(df$date, na.rm = TRUE),
                                        to = max(df$date, na.rm = TRUE), by = 1)) %>%
    mutate(day_number = as.numeric(difftime(date, min(df$date, na.rm = TRUE) - 1)))
  df.2 <- df.1 %>%
    full_join(full.date.df, by = "date") %>% arrange(date, site)
  return(df.2)
}
# To get leaf fall on each day of the interval
leaf.fall.daygaps <- lapply(df.sp_site, fill.day.gaps)

@@

6. New: Using Joe's suggestion: Could you estimate living canopy leaf mass per unit ground area present at time t (L[t]) from mean leaf longevity and leaf mass fall per unit ground area at time t (F[t]) as follows:  L[t] = F[t+1] + F[t+2] + … + F[t + mean leaf longevity]

@@

```{r leaf-fall-LAI, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{leaf-fall-LAI}Seasonality curves for Leaf Fall in LAI units based on Joe's calculation. LL = species-specific leaf lifespan in days. Species are color coded by deciduousness.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/Leaf_fall_LAI.seasonality_mean.jpeg"))
```

```{r LAI-ssn-loop, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{LAI-ssn-loop}Mean LAI seasonality curves based on Joe's approach to calculation. LL = species-specific leaf lifespan in days. Species are color coded by deciduousness.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/LAI.seasonality_mean_loop.jpeg"))
```

```{r LAI-ssn-loop-subcol, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{LAI-ssn-loop-subcol}Mean LAI seasonality curves based on Joe's approach to calculation. LL = species-specific leaf lifespan in days. Species are color coded by the method/source of leaf-lifespan substitution.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/LAI.seasonality_mean_loop_color_substitution.jpeg"))
```

>>>>

1. Gap-filled species-level mean leaf lifespan with observations from crane sites (data referred to as "cohort""); if not with genus-level mean; else family-level mean. When neither were available: 
    1. For deciduous species, used mean lifespan at PNM site
    2. For evergreen species, used a relationship of leaf lifespan (LL) with LMA. This LMA-based relationship was based on data at San-Lorenzo (FTS) for canopy leaves and for species with sufficient sample sizes (n_lifespan > 100) (Figure \@ref(fig:ll-lma)).

    Per Joe, BCI climate is more similar to San Lorenzo, while PNM site is much drier. More importantly, the BCI 50-ha plot and San Lorenzo support old-growth forests. PNM supports much younger, secondary forest (age unknown, perhaps 100 years). So, used the leaf lifespan-LMA relationship for San Lorenzo. This was used to predict leaf lifespan data for species without leaf lifespan data based on their LMA, after gap-filling species LMA data:  
    Gap-filled LMALAM_AVD with the following sequence of priority:    
    LMALAM_AVD relationship with LMADISC_AVD > LMALAM_AVD rel. with LMALEAF_AVD > Genus-level mean LMALAM_AVD (gap-filled with LMADISC_AVD/LMALEAF_AVD) > Family-level mean LMALAM_AVD (gap-filled with LMADISC_AVD/LMALEAF_AVD)

@@

fill.day.gaps <- function(df) {
  df.1 <- df %>% arrange(date) %>%
    # for the first census assume leaf fall is collected over the past week
    mutate(n.days = c(7, as.numeric(date - lag(date))[-1])) %>%
    #  leaf_fall rate per day for the census interval is stored on the day of the census:
    mutate(leaf_gm_rate = leaf_gm_per_m2/n.days,
           # approx() in the next function will interpolate this forward until the next census,
           # we want it to interpolate this value backward over the past census interval,
           # so just for computation purposes, leading this value
           leaf_gm_rate_lead = lead(leaf_gm_rate, n = 1))
  # Gap-filling dates
  full.date.df <- data.frame(date = seq(from = min(df$date, na.rm = TRUE),
                                        to = max(df$date, na.rm = TRUE), by = 1)) %>%
    mutate(day_number = as.numeric(difftime(date, min(df$date, na.rm = TRUE) - 1)))
  df.2 <- df.1 %>%
    full_join(full.date.df, by = "date") %>% arrange(date, site)
  return(df.2)
}
# To get leaf fall on each day of the interval
leaf.fall.daygaps <- lapply(df.sp_site, fill.day.gaps)

@@

6. Joe's suggestion: Estimate living canopy leaf mass per unit ground area present at time t (L[t]) from mean leaf longevity and leaf mass fall per unit ground area at time t (F[t]) as follows:  L[t] = F[t+1] + F[t+2] + … + F[t + mean leaf longevity]

@@

```{r leaf-fall-LAI, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{leaf-fall-LAI}Seasonality curves for Leaf Fall in LAI units. LL = species-specific leaf lifespan in days. Species are color coded by deciduousness.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/Leaf_fall_LAI.seasonality_mean.jpeg"))
```

```{r LAI-ssn-loop, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{LAI-ssn-loop}Mean LAI seasonality curves. LL = species-specific leaf lifespan in days. Species are color coded by deciduousness.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/LAI.seasonality_mean_loop.jpeg"))
```

```{r LAI-ssn-loop-subcol, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{LAI-ssn-loop-subcol}Mean LAI seasonality curves with species color coded by the source of leaf-lifespan observation/substitution. LL = species-specific leaf lifespan in days.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/LAI.seasonality_mean_loop_color_substitution.jpeg"))
```
